---
title: "Plot Beta Parameters from Fraser Sockeye Data"
author: "Carrie Holt"
date: "2023-07-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Plot beta parameters from Fraser Sockeye Data

Parameter estimates are from Ann-Marie Huang (DFO, pers. comm. 2022). Parameters were estimated in 2021 for Fraser Sockeye Salmon populations using a Larkin stock-recruitment model

```{r estimate}
# Libraries
library(dplyr)
library(ggplot2)
library(cowplot)
source(here::here("report/ppnErr.R"))

# Read in parameter estimates
pars <- read.csv(file = here::here("report/Larkin-summary-stats_sd2021_feb.csv"))

df <- pars %>% filter(X!="alpha"&X!="deviance"&X!="sigma") %>% 
	dplyr::select(X, stk.name, mean) %>% 
	group_by(stk.name) %>% 
	mutate(mBeta = mean(mean), sBeta = sum(mean)) %>% 
	ungroup() %>% 
	mutate(rel= mean/mBeta) %>% 
	mutate(ppn = mean/sBeta)

alpha <- pars %>% filter(X=="alpha") %>% 
	dplyr::select(X, stk.name, mean) %>% 
	mutate(alpha = exp(mean)) %>% 
	select(!mean)
#summary(alpha$alpha)
sigma <- pars %>% filter(X=="sigma") %>% 
	dplyr::select(X, stk.name, mean)
#summary(sigma$mean)

```

## Plot beta parameters

Plot beta scaled parmeters (sum to one), grouped into three categories.

```{r plot, echo=FALSE}

ggplot(df, aes(x=X, y=ppn, group = stk.name)) + geom_line() + 
	facet_wrap(vars(as.factor(stk.name)))#geom_line(aes(colour=stk.name))

df.LargeB0 <- df %>% filter(stk.name == "Birkenhead"|
													stk.name == "Bowron"|
													stk.name == "Cultus"|
													stk.name == "Fennel") # maybe Raft?

p1 <- ggplot(df.LargeB0, aes(x=X, y=ppn, group = stk.name)) + geom_line() + 
	geom_line(aes(colour=stk.name)) + ggtitle("Stocks with large beta0") +
	ylim(0,0.6)

df.ConstantB0 <- df %>% filter(stk.name == "Gates"|
															stk.name == "L. Shuswap"|
															stk.name == "Nadina"|
															stk.name == "Quesnel"|
																stk.name == "Seymour"|
																stk.name == "Weaver") # maybe Raft?

p2 <- ggplot(df.ConstantB0, aes(x=X, y=ppn, group = stk.name)) + geom_line() + 
	geom_line(aes(colour=stk.name)) + 
	ggtitle("Stocks with relative similar betas") + ylim(0,0.6)

df.VariableB0 <- df %>% filter(stk.name == "Chiko"|
															 	stk.name == "E. Stuart"|
															 	stk.name == "Harrison"|
															 	stk.name == "L. Stuart"|
															 	stk.name == "Portage"|
															 	stk.name == "Raft"|
															 	stk.name == "Scotch"|
															 	stk.name == "Stellako"|
															 	stk.name == "Upper Pitt") # maybe Raft?

p3 <- ggplot(df.VariableB0, aes(x=X, y=ppn, group = stk.name)) + geom_line() + 
	geom_line(aes(colour=stk.name)) + 
	ggtitle("Stocks with variable betas") + ylim(0,0.6)

plot_grid(p1, p2, p3, nrow= 2, ncol=2)

```
 
 Plot simlated beta parameters for each of those three categories 
```{r siulatedPlots}

gen.ppns <- function(nTrials, scenario, err){
	MCtrial <- 1
	sim <- ppnErr(scenario, err, runif(4,0,1))
	df.sim <- data.frame(X = c("beta0", "beta1", "beta2", "beta3"), 
											 MCtrial, 
											 ppn=sim)
	for (i in 2:nTrials){
		sim <-  ppnErr(scenario, err, runif(4,0,1))
		if (round(sum(sim), 2) != 1) print(cat("Warning proportions do sum to 1"))
		MCtrial <- i
		df.sim <- add_row(df.sim, 
											data.frame(X = c("beta0", "beta1", "beta2", "beta3"), 
																 MCtrial, 
																 ppn=sim))
	}
	return(df.sim)
}

# Simulated betas, where beta0 is large (group 1)
set.seed(10)
scenario <- c(0.45,0.25,0.15,0.15)
nTrials <- 100
err <- 0.3
df.sim <- gen.ppns(nTrials, scenario, err)

ggplot(df.sim, aes(x=X, y=ppn, group = as.factor(MCtrial))) +
	geom_line() +
	geom_line(aes(colour=as.factor(MCtrial))) +
	guides(color = FALSE,  scale = "none") + ylim(0,0.6)


# Simulated betas, where bete values are even with low variability (group 2)
df.sim <- gen.ppns(nTrials=100, scenario = rep(0.25,4), err = 0.2)

ggplot(df.sim, aes(x=X, y=ppn, group = as.factor(MCtrial))) +
	geom_line() +
	geom_line(aes(colour=as.factor(MCtrial))) +
	guides(color = FALSE,  scale = "none") + ylim(0,0.6)


# Simulated betas, where beta values are variable, but no trend (group 3)
df.sim <- gen.ppns(nTrials=100, scenario = rep(0.25,4), err = 0.6)

ggplot(df.sim, aes(x=X, y=ppn, group = as.factor(MCtrial))) +
	geom_line() +
	geom_line(aes(colour=as.factor(MCtrial))) +
	guides(color = FALSE,  scale = "none") + ylim(0,0.6)

# 
# # BASE CASE PARS
# x <- salmon_run(alpha = 7,
# 								beta = rep(0.25,4),
# 								p_prime = c(0.003, 0.917, 0.080),
# 								rho = 0,
# 								omega = 0.4,
# 								sigma_nu = 1,
# 								phi_1 = 0,
# 								T = 100,
# 								h_t = NULL,
# 								R_t_init = c(25, 5, 1, 1, 25, 5, 1, 1)*0.05,
# 								deterministic = FALSE,
# 								extirp = 2e-5)
# 

```
 
 