---
title: "plotRealisations"
author: "Carrie Holt"
date: "2023-07-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Plot realisations

This code runs the sim_and_fit_realisations over numerous trials and extracts the forecasts

```{r runRealisations}

library(dplyr)
library(ggplot2)
library(larkin)

# Run using stan-optim for Larkin model to speed up model runs
# See larkin-compareMLE-Bayes.Rmd for a comparison
# larkin_args() and ricker_args() have default values, but if one element 
# changes, the entire list needs to be included here
run <- EDMsimulate::sim_and_fit_realisations(salmon_sim_args = 
																						 	list(sigma_nu = 1, 
																						 			 omega = 0.4), 
																						 larkin_args = list(
																						 	run_stan = FALSE,
																						  prior_mean_alpha = 2,
																						  prior_mean_beta = -rep(1,4),
																						  prior_mean_sigma = 0.5,
																						  prior_sd_alpha = 0.5,
																						  prior_sd_beta = rep(0.25,4),
																						  prior_sd_sigma = 0.25),
																						 ricker_args = list(
																						 	run_stan = FALSE,
																						 	prior_mean_alpha = 2,
																						 	prior_mean_beta = -1,
																						 	prior_mean_sigma = 0.5,
																						 	prior_sd_alpha = 0.5,
																						 	prior_sd_beta = 0.25,
																						 	prior_sd_sigma = 0.25),
																						 M = 100)
```
Then plot outputs.

```{r plotForecasts}

run <- run %>% dplyr::mutate(absErr_edm = 
														 	(R_prime_T_edm_fit - R_prime_T_sim), 
														 absErr_lar = 
														 	(R_prime_T_lar_fit - R_prime_T_sim),
														 absErr_ric = 
														 	(R_prime_T_ric_fit - R_prime_T_sim))

absErr <- tidyr::pivot_longer(run, c(absErr_edm, absErr_lar, absErr_ric), 
																names_to="forMethod", 
																names_prefix = "absErr_", 
																values_to="absErr")
absErr <- absErr %>% select(m,forMethod, absErr)

forecast_data <- tidyr::pivot_longer(run, c(R_prime_T_edm_fit, 
																						R_prime_T_lar_fit,
																						R_prime_T_ric_fit), 
															names_to="forMethod", 
															names_prefix = "R_prime_T_", 
															values_to="R_prime_T_fit") 
forecast_data <- forecast_data %>% 
	select(m,forMethod, R_prime_T_fit, R_prime_T_sim)

p1 <- forecast_data %>% ggplot(aes(x=R_prime_T_sim, 
																	 y=R_prime_T_fit, 
																	 colour=forMethod)) + 
	geom_point() + 
	geom_smooth(method='lm', formula= y~x) + 
	xlab("Simulated 'true' recruitment") + 
	ylab("Forecasted recruitment") +
	geom_abline(slope=1, intercept = 0, col="dark grey")
p1
p2 <- absErr %>% ggplot(aes(x=absErr, colour=forMethod)) + 
	geom_density() +
	xlab("Absolute error in forecast") +
	ylab("Density")
p2

```
The Larkin and Ricker models tend to have absolute errors in forecast that are closer to zero, compared with EDM.

The simulation is then repeated, but with altered beta parameters where beta0 is dominant over the remaining beta parameters (more similar to a Ricker model).

```{r Beta0}
run2 <- EDMsimulate::sim_and_fit_realisations(salmon_sim_args = 
																						 	list(sigma_nu = 1, 
																						 			 omega = 0.4,
																						 			 beta = c(0.45, 0.25, 0.15, 0.15)), 
																						 larkin_args = list(
																						 	run_stan = FALSE,
																						  prior_mean_alpha = 2,
																						  prior_mean_beta = -rep(1,4),
																						  prior_mean_sigma = 0.5,
																						  prior_sd_alpha = 0.5,
																						  prior_sd_beta = rep(0.25,4),
																						  prior_sd_sigma = 0.25),
																						 ricker_args = list(
																						 	run_stan = FALSE,
																						 	prior_mean_alpha = 2,
																						 	prior_mean_beta = -1,
																						 	prior_mean_sigma = 0.5,
																						 	prior_sd_alpha = 0.5,
																						 	prior_sd_beta = 0.25,
																						 	prior_sd_sigma = 0.25),
																						 M = 100)


run2 <- run2 %>% dplyr::mutate(absErr_edm = 
														 	(R_prime_T_edm_fit - R_prime_T_sim), 
														 absErr_lar = 
														 	(R_prime_T_lar_fit - R_prime_T_sim),
														 absErr_ric = 
														 	(R_prime_T_ric_fit - R_prime_T_sim))

absErr2 <- tidyr::pivot_longer(run2, c(absErr_edm, absErr_lar, absErr_ric), 
																names_to="forMethod", 
																names_prefix = "absErr_", 
																values_to="absErr")
absErr2 <- absErr2 %>% select(m,forMethod, absErr)

forecast_data2 <- tidyr::pivot_longer(run2, c(R_prime_T_edm_fit, 
																						R_prime_T_lar_fit,
																						R_prime_T_ric_fit), 
															names_to="forMethod", 
															names_prefix = "R_prime_T_", 
															values_to="R_prime_T_fit") 
forecast_data2 <- forecast_data2 %>% 
	select(m,forMethod, R_prime_T_fit, R_prime_T_sim)

p3 <- forecast_data2 %>% ggplot(aes(x=R_prime_T_sim, 
																	 y=R_prime_T_fit, 
																	 colour=forMethod)) + 
	geom_point() + 
	geom_smooth(method='lm', formula= y~x) + 
	xlab("Simulated 'true' recruitment") + 
	ylab("Forecasted recruitment") +
	geom_abline(slope=1, intercept = 0, col="dark grey")
p3
p4 <- absErr2 %>% ggplot(aes(x=absErr, colour=forMethod)) + 
	geom_density() +
	xlab("Absolute error in forecast") +
	ylab("Density")
p4

```

