---
title: "salmon_sim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{salmon_sim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  fig.width = 6
)
```

Different ways to simulate data, and to fit it, and to do multiple simulations
with different stochasticity or assumptions/scenarios. Haven't yet decided on
exact terminology.
The notation in the code matches that in the write up of the second manuscript (in the `edm-work` repository).

```{r setup}
# load_all()          # use this for local up-to-date testing
library(EDMsimulate)
library(pbsEDM)      # https://github.com/pbs-assess/pbsEDM
```

## To run a simulation
```{r}
sim_data <- salmon_sim()
sim_data
```

##  Simple plot of resulting time series of spawners
```{r}
plot_sim(sim_data)
```

## To run and plot in one go
```{r}
salmon_run()
```

See `?salmon_sim` for settings to change the default parameters and initial
conditions for the simulations.

## To do bifurcation diagrams

See code in `salmon_bif.R`, not needed here yet.

## To fit using pbsEDM

Basing on example analysis in pbsEDM vignette `analyse_simple_time_series.Rmd`.

```{r fit1}
sim_fit <- pbsEDM::pbsEDM(sim_data,
                          lags = list(R_prime_t = 0,     # can test
                                        # others. Biologically defendible assumptions
                                      S_t = 0:3),
                          first_difference = TRUE)    # can try both
sim_fit$results
```

Try a deterministic simulation, and then again using the new `sim_and_fit()`
wrapper:
```{r function}
sim_data_determ <- salmon_sim(deterministic = TRUE)

sim_fit_determ_fit <- pbsEDM::pbsEDM(sim_data_determ,
                          lags = list(R_prime_t = 0,     # can test
                                        # others. Biologically defendible assumptions
                                      S_t = 0:3),
                          first_difference = TRUE)    # can try both

# Then in one function
sim_fit_determ_fit_2 <- sim_and_fit(salmon_sim_args = list(deterministic =
                                                             TRUE),
                                    pbsEDM_args = list(lags = list(R_prime_t = 0,
                                                                   S_t = 0:3),
                                                       first_difference = TRUE))

# These should match:
sim_fit_determ_fit$results
sim_fit_determ_fit_2$fit$results
expect_equal(sim_fit_determ_fit$results,
             sim_fit_determ_fit_2$fit$results)
```

## To run multiple realisations using the same parameter values but differing in stochasticity

Using the term `realisations` specifically for multiple runs that
have the same parameters but only differ in the resulting process noise
(i.e. just different due to stochasticity). Then `scenarios` for
different assumptions of parameter values. Maybe scenarios A, B, and C for the
three sets of assumptions on
the proportions that we talked about last week, since they're the main things we
want to change, and then numbered scenarios for changing other things (and A1,
A2, ..., for combining them).


TODO centre and scale for all fits, go back and do functions and tests.

Then vary over E, but not simple as need to vary over multivariate lags. So need
to pick the best combination of lags, not the best E.

Then do s-map as well using the best lags? May as well, to get the best
results. Then do the uncertainty ideas on just the optimal ones (getting a
confidence interval for each realisation).
